---
title: "Untitled"
author: "Tim Lamour"
date: "11/22/2020"
output: github_document
---

# Intro

Explore processed EEMB 144L 16s sequences using phyloseq


```{r}
library(tidyverse)
library(phyloseq)
library(RColorBrewer)
```

# Import Data

```{r}
count.tab <- read_rds("~/Desktop/Github/144l_students/Input_Data/week 6/seqtab-nochimtaxa.rds")
tax.tab <- read_rds("~/Desktop/Github/144l_students/Input_Data/week 6/taxa.rds")
sample.tab <- read_rds("~/Desktop/Github/144l_students/Output_data/Week 4/144L_processed_DOC_BGE.rds") %>%
  drop_na(DNA_SampleID) %>%
  column_to_rownames(var = "DNA_SampleID")
```

# Phyloseq Object 

```{r}
OTU = otu_table(count.tab, taxa_are_rows = TRUE)
TAX = tax_table(tax.tab)
SAM = sample_data(sample.tab)
ps = phyloseq(OTU,TAX,SAM)
```

# Filter Sequences 

filter out chloroplassts and mitochondria since we only intended to amplify bacterial sequences

```{r}
sub_ps <- ps %>%
  subset_taxa(Family != "mitochondira" & Order != "Chloroplast")
```


# Sample Summary 

As a first analysis, we will look at the distribution of read counts from our samples

```{r}
sample_sum_df <- data.frame(sum = sample_sums(sub_ps))
```

```{r}
# mean, max and min of sample read counts 
summary(sample_sum_df)
```

```{r}
sub_ps <- ps %>%
  subset_taxa(Family != "mitochondira" & Order != "Chloroplast")
ggplot(sample_sum_df, aes(x = sum)) +
  geom_histogram(color = "black", fill = "#377EB8", binwidth = 5000) +
  ggtitle("Distribution of sample sequence depth") +
  xlab("Read counts") +
  theme(axis.title.y = element_blank()) +
  theme_bw()
```


# Beta Diversity

beta diversity = patterns in community strucutre between samples 

## Subsample 

we will rarefy (random subsample with replacement) the read depth of the samples first (scale to the smallest library size)

```{r}
ps_min <- rarefy_even_depth(sub_ps, sample.size = min(sample_sums(sub_ps)))

mean(sample_sums(sub_ps)) #28796.92
mean(sample_sums(ps_min)) #2064
```

## NMDS 

Here we will look at non-metric multidimensional scaling (NMDS) ordinations of our full community samples. For NMDS plots it's important to set a seed since the starting positions of samples in the algorithm is random 

```{r}
set.seed(1)
# Ordinate
nmds <- ordinate(sub_ps, method = "NMDS", distance = "bray") # stress = 0.07
```

```{r}
set.seed(1)
# Ordinate 
nmds_min <- ordinate(ps_min, method = "NMDS", distance = "bray") # stress = 0.08
```

```{r}
nmds.plot <- plot_ordination(sub_ps, nmds, title = "NMDS") +
  geom_point(aes(fill = days, shape = factor(Treatment)), alpha = 0.6, stroke = 2, size = 4) + 
  scale_shape_manual(values = c(21, 22, 23, 24)) +
  scale_fill_gradient(low = "#0db5e6", high = "#d31f2a") +
  theme_bw()


#removing one of the plotting layers (there are points within points)
nmds.plot$layers <- nmds.plot$layers[-1]
nmds.plot +
  facet_grid(~Location) +
  guides(fill = guide_colorbar(title = "Days"), shape = guide_legend(title = "Treatment"))

```

```{r}
nmds_min.plot <- plot_ordination(ps_min, nmds_min, title = "NMDS") +
  geom_point(aes(fill = days, shape = factor(Treatment)), alpha = 0.6, stroke = 2, size = 4) +
  scale_shape_manual(values = c(21, 22, 23, 24)) +
  scale_fill_gradient(low = "#0db5e6", high = "#d31f2a") +
  theme_bw()

nmds_min.plot$layers <- nmds_min.plot$layers[-1]
nmds_min.plot +
  facet_grid(~Location) +
  guides(fill = guide_colorbar(title = "Days"), shape = guide_legend(title = "Treatment"))

```

These graphs show that the patterns between sub-sampling and not sub-sampling are essentially the same, and that we can go forward with our standardized data set. The stress of this plot was ~0.08 so this is an accurate representation of this data. 


# Alpha Diversity 

we are going to calculate the Chao1 index for richness and the Shannon diversity index. We will use the subsampled library, which retains estimates of the species abundance of the real population while standardizing sampling effort. 

```{r}
richness <- estimate_richness(ps_min, measures = c("Chao1", "Shannon")) %>%
  rownames_to_column(., var = "DNA_ID") %>%
  mutate_at(vars(DNA_ID), str_replace_all, pattern = "X144", "144")
```


let's add the sample metadata into this dataframe 

```{r}
alphadiv <- left_join(richness, sample.tab %>%
  rownames_to_column(., var = "DNA_ID"))
```

```{r}
library(ggpubr)
```


```{r fig.height=8, fig.width=12}
pivot.data <- alphadiv %>%
  select(Treatment, Location, Bottle, Timepoint, days, Chao1, Shannon) %>%
  pivot_longer(., cols = c(Chao1, Shannon), names_to = "measure", values_to = "est") %>%
  left_join(., alphadiv %>%
              select(Treatment, Location, Bottle, Timepoint, days, se.chao1)) %>%
  mutate(se.chao1 = ifelse(measure == "Chao1", se.chao1, NA))


alpha.plot <- ggboxplot(pivot.data, x = "Timepoint", y = "est", 
                        xlab = expression(italic(paste(""))),
                        ylab = expression(italic(paste("Alpha Diversity Measure"))),
                        add = "dotplot", 
                        width = 0.2,
                        ggtheme = theme_bw()) +
  stat_compare_means(label.x = "6") +
  facet_grid(measure~ factor(Treatment), scales = "free")
  

```

## Generate relative abundances

```{r}
ps_std <- transform_sample_counts(ps_min, function(x) x/sum(x))
ps_std.tab <- as(otu_table(ps_std), "matrix")
ps_std.df = as.data.frame(ps_std.tab)
```

## Make table 


```{r}
tax.df <- as.data.frame(tax.tab)
custom.tab <- tax.df %>%
  rownames_to_column(., var = "asv") %>%
  left_join(., ps_std.df %>% rownames_to_column(., var = "asv")) %>%
  mutate(pco = paste(Phylum, "_", Class, "_", Order)) %>%
  select(-c(asv:Genus)) %>% 
  select(pco,everything()) %>%
  group_by(pco) %>%
  summarise_at(vars(contains(c("144"))), sum, na.rm = T) %>% 
  ungroup()

#save the row names and then  make them into the column names 
colnames <- custom.tab[,1]
t_custom.tab <- as.data.frame(t(custom.tab[,-1]))
colnames(t_custom.tab) <- colnames$pco

#merge
sweet.tab <- t_custom.tab %>%
  rownames_to_column(., var = "sample") %>%
  left_join(., sample.tab %>% rownames_to_column(., var = "sample") %>% select(sample, Experiment, Location, Bottle, Treatment, Timepoint, days, cells)) %>%
  select(sample, Experiment:cells, everything())

relabund <- sweet.tab %>%
  select(-c(sample:cells)) %>%
  .[ , colSums(.) > 0] %>%
  .[, order(colSums(-.))] %>%
  bind_cols(sweet.tab %>% select(sample:cells), .)

```

## Heatmap 

```{r fig.height=10, fig.width=8}
relaheat.data <- relabund %>%
  select(-c(sample, Experiment, Location, Bottle, days, cells)) %>%
           pivot_longer(.,-c(Treatment:Timepoint), names_to = "taxa", values_to = "relabund") %>%
           separate(taxa, into = c("p", "c", "o"), sep = "_")

library(viridis)

relabund <- relaheat.data %>%
  ggplot(aes(x = Timepoint, y = o)) +
  geom_tile(aes(fill = relabund), color = "white") +
  scale_fill_viridis(option = "D") +
  labs(x = "", y = "Order", fill = "Relative Abundance") +
  facet_grid(~factor(Treatment)) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 12),
        legend.position = "top") +
  guides(fill = guide_colourbar(barheight = 2, barwidth = 20, frame.colour = "black", frame.linewidth = 2, ticks.colour = "black", ticks.linewidth = 1), color = F)

```

This figure shows that the bacterial community composition of the different treatments changes over time, and that there are different communities present in the different treatments. For example, the GNP treatment becomes almost completely dominated by Oceanospirillales. The other treatments all do see a rise in Oceanospirillales, but not nearly to the same extent as GNP. Additionally, some orders are present in some treatments but are hardly present in others. For example, campylobacterales are very present in the ash lechate but are not in the others. This could indicate that there is some nutrient source in each one of these treatments that favors the growth of a barticular type of bacteria. 


# Save and knit 

```{r}
saveRDS(sweet.tab, "~/Desktop/Github/144l_students/Output_data/Week 6/Custom_ASV_Table.rds")

saveRDS(sub_ps, "~/Desktop/Github/144l_students/Output_data/Week 6/phyloseq_obj.rds")

saveRDS(ps_min, "~/Desktop/Github/144l_students/Output_data/Week 6/subsampled_phyloseq_obj.rds")

saveRDS(alphadiv, "~/Desktop/Github/144l_students/Output_data/Week 6/alphadiv.rds")
```











